A Framework for Equivalence of Cast Calculi and a new Cast Representation

### 2020-10-05 paper.tex
Updates from Overleaf

%% \todo[inline]{KC: Cite "Contracts for higher-order functions" for the proxy-based approach maybe?}
%% \todo[inline]{JS: That paper doesn't exactly fit into what you're talking about in the above
%% sentence. It might be appopriate to cite it elsewhere in the paper. Also,
%% there are earlier papers about coercive subtyping (Computing with coercions by Breazu-Tannen et al)
%% that use a proxy-based approach. The contracts paper may be the first to propose blame tracking.}

@@

\todo[inline]{JS: say something about the difference between the above
  and what is in this paper. That's the whole point of the related work
  section from the eyes of the reviewer.}

@@

Fig.~\ref{fig:R-machine-syntax} defines machine 
states $s$, including a definition of values $v$ and value typing.
Fig.~\ref{fig:R-machine-dynamics} defines the transition relation of the CEK
machine (\CMachine{B}).
We name the machine after \textbf{c}ast calculus.
%
This machine is parameterized over blame strategy $B$, which ranges over
Lazy D (Fig.~\ref{fig:blame-strategy-LazyD}) and Lazy UD 
(Fig.~\ref{fig:blame-strategy-LazyUD}).
\todo[inline]{JS: I don't think $B$ is restricted to just those two.
Can't it be any blame strategy that satisfies certain conditions?}
\todo[inline]{KC: Technically a $B$ should define a predicate over pretypes, which specifies
what is injectable, and a apply-cast function. Of course, the set of values that the apply-cast
function works with depends on the set of injectable types.
\url{https://github.com/Gradual-Typing/Equivalence-of-Cast-Calculi/blob/master/equivalence-of-cast-calculi/R/BlameStrategies.agda}
Do you want me to put these contraints in the paper?}
\todo[inline]{JS: Yes.}
%
The transitions involving casts are highlighted in red and described in
more detail below. The other transitions are standard for a CEK
machine for an extended simply typed lambda calculus~\citep{Felleisen:2009aa}.
%
Recall that CEK machines involve two kinds of transitions: those that
dive further into an expression (looking for a redex) and push an
entry onto the continuation and those that return a value to the
current continuation and possibly perform a computation.  That is, the
machine state is either in an evaluating $\sexpr{e}{E}{K}$ or
returning $\scont{v}{K}$ configuration, respectively. Additionally,
there are two halting configurations: \shalt{v} halts with a value,
while \error{l} halts with a blame label.  Observables include all the
value constructors and blame labels, as in the work of
\citet{siek2012interpretations}.  The function that converts values to
observables ($observe(v) = o$) is defined in the obvious way.

Let $v$ range over values. The $\vdyn{I}{v}$ stands for values casted
to the dynamic type, as in the work of \citet{wadler2009well}. The
inner value must be of an injectable type. The definition of
injectable types depends on blame strategies:
%
for the \LD\ strategy, every pre-type is injectable 
(Fig.~\ref{fig:blame-strategy-LazyD});
%
for the \LUD\ strategy, only ground types are injectable 
(Fig.~\ref{fig:blame-strategy-LazyUD}).
%
Other kinds of values are either Boolean constants, functions, casted 
functions, pairs, or casted pairs.
%
The \proxy{v}{\ccast{S}{l}{T}} partial function applies the cast to $v$, which must
have type $S$. This function is handy when we describe the dynamic semantics.
% In earlier version of this work, 
% we were tempted to use a generic constructor for all casted value, and use 
% value typing rules to restrict the cast such that its source and target are 
% shallowly consistent pre-types. This approach unfortunately considers 
% $\ecast{\etrue}{\ccast{\Pbool}{l}{\Pbool}}$ as a value.

The syntax \error{l} occurs several times in the syntax definition. We shall 
view those nonterminals that have an \error{l} case as error monads. We use 
the symbol \mbind{} as the monad bind operator, which is defined as follows:
\[
\begin{array}{rcll}
(\error{l}) \mbind f & = & \error{l} \\
          a \mbind f & = & f(a)      & \text{otherwise}
\end{array}
\]
\todo[inline]{JS: why not use monadic let instead? That avoids
  the need to also use meta-lambdas, which you would need to
  explain, e.g., in the transition rule that uses applyCast.}

@@

The transition relation \reduceto{s}{s} invokes the cast application
function, noted as $\applyCast{v}{\ccast{S}{l}{T}}$ to capture the
differences between blame strategies
(Fig.~\ref{fig:blame-strategy-LazyD} and
Fig.~\ref{fig:blame-strategy-LazyUD}).
\todo[inline]{JS: be consistent with the font used for applyCast. Sometimes
  you use roman and sometimes you used italic. Also, be consistent
  for all the metafunctions (eval, observe, etc.).}
%
When evaluating a cast expression, the machine moves the cast to the
continuation and evaluates the inner expression.
%
To cast a value, the machine invokes applyCast on the value. The cast
result is either a value or an error (\error{l}). If the cast
succeeds, the machine returns the value to the next continuation.  If
the cast fails, the machine halts with the blame label.
%
To apply a casted function, the machine first casts the argument $v$, then
applies the inner function $u$ to the casted argument, and finally
casts the return value.
%
To project the first element of a casted pair, the machine projects
the first element of the inner pair $u$ and cast it.
%
Projecting the second element of a casted pair is similar.

@@

@@

In this section, we review the coercions
of~\citet{henglein1994dynamic} and \citet{herman2010space} and the
normal form of \citet{siek2012interpretations} to motivate the design
of hypercoercions.  We omit sum types and product types in this
section because the prior work did not include them.
\todo[inline]{But Fig 6 includes product types.}

@@

Coercions come with a reduction relation so it is natural to ask about
their normal forms. \figref{fig:normal-coercion} defines the syntax of
coercion in normal form, written $\hat{c}$.
\todo[inline]{JS: In the figure, normal forms are written $s$ and $t$.}
Conceptually, a coercion
in normal form has three parts, all of which are optional.  It may
start with a projection $I?^l$, followed by a function coercion, and
then concluded with an injection $I!$ or a failure $\bot^l$. While
this is a simple idea, there are 10 clauses in the definition for
$\hat{c}$!

@@

% All cast representation shall be proved correct, that is, besides reducing 
% space consumption, using a cast representation should not affect the results of 
% program execution.
In this section we present a framework for proving the 
correctness of space-efficient cast representations.
The framework includes an abstract data 
type named Cast ADT, which specifies the operators of cast representations, 
a CEK machine $\SMachine{C}$ parameterized over a Cast ADT,
which uses the cast operators implemented by $C$ to work with casts, 
and theorems relating $\SMachine{C}$ to $\CMachine{B}$.
%
Subsection~\ref{sec:cast-adt-and-space-efficient-cek} describes Cast ADT and the 
machine \SMachine{C}. We conjecture that the machine \SMachine{C} is 
space-efficient provided that the cast representation $C$ performs compression.
%
Subsection~\ref{sec:good-cast-adt} defines two subset of Cast ADT, namely Lazy D 
Cast ADT and Lazy UD Cast ADT. As their names suggest, if $C$ is a Lazy D Cast 
ADT, executing a program in $\SMachine{C}$ would give the same result as in 
$\LDMachine$. Similarly, if $C$ is a Lazy UD Cast ADT, executing a program in 
$\SMachine{C}$ would give the same result as in $\LUDMachine$.
%
We prove the theorems relating \SMachine{C} and \CMachine{B} in
Subsection~\ref{sec:S-correct}.
The framework is written in the Agda language~\citep{norell2008dependently}.

>>>>

\todo[inline]{KC: Cite "Contracts for higher-order functions" for the proxy-based approach maybe?}
\todo[inline]{JS: That paper doesn't exactly fit into what you're talking about in the above
sentence. It might be appopriate to cite it elsewhere in the paper. Also,
there are earlier papers about coercive subtyping (Computing with coercions by Breazu-Tannen et al)
that use a proxy-based approach. The contracts paper may be the first to propose blame tracking.}
\todo[inline]{KC: I think readers should be able to trace back history by following the pointers in the 
intro section. I don't know whether my thought is correct. I would rather not giving misleading
pointers. I also don't know how far should a pointer chases back. Readers can be directed to recently
works which will eventually lead them back to the earliest possible paper. But wouldn't it be
great if the author of a new paper shows reader all the milestones? After all we already know 
pointer dereferencing is expensive.}

@@

\todo[inline]{JS: say something about the difference between the above
  and what is in this paper. That's the whole point of the related work
  section from the eyes of the reviewer.}
\todo[inline]{KC: I might have some misconcept about the related work
section. You might have felt this, haha. Now I think this section is
targeting the following scenarios:
- ``I think someone has solved this problem ...''
- ``This work looks similar to paper X, what is the difference?''
- ``Is this work inspired by paper X?''
Is my understanding correct?}

@@

Fig.~\ref{fig:R-machine-syntax} defines machine 
states $s$, including a definition of values $v$ and value typing.
Fig.~\ref{fig:R-machine-dynamics} defines the transition relation of the CEK
machine (\CMachine{B}).
We name the machine after \textbf{c}ast calculus.
\todo[inline]{JS: explain the name choice of $\mathcal{R}$.}
\todo[inline]{KC: I want to space-efficient machine to be S, so ... B might be a better name 
because it reminds people of the $\lambda B$ calculus.
Or $T$ for type-based cast?
Or $C$ for cast-calculus?
Or $D$ for definitional semantic?
How do you think?}
\todo[inline]{JS: B or C both sound like good choices for the reasons
you give.}
\todo[inline]{KC: Let's go with C to avoid confusion with Blame strategy.}
%
% This machine is parameterized over blame strategy $B$, which ranges over
% Lazy D (Fig.~\ref{fig:blame-strategy-LazyD}) and Lazy UD 
% (Fig.~\ref{fig:blame-strategy-LazyUD}).
This machine is parameterized over blame strategy $B$. We will describe
the components of a blame strategy as we go through the machine definition.
In this work we focus on Lazy D (Fig.~\ref{fig:blame-strategy-LazyD}) and Lazy UD 
(Fig.~\ref{fig:blame-strategy-LazyUD}). Adding more blame strategies, 
such as the ones defined in \cite{siek2009exploring}, is possible.
\todo[inline]{JS: I don't think $B$ is restricted to just those two.
Can't it be any blame strategy that satisfies certain conditions?}
\todo[inline]{KC: Technically a $B$ should define a predicate over pretypes, which specifies
what is injectable, and a apply-cast function. Of course, the set of values that the apply-cast
function works with depends on the set of injectable types.
\url{https://github.com/Gradual-Typing/Equivalence-of-Cast-Calculi/blob/master/equivalence-of-cast-calculi/R/BlameStrategies.agda}
Do you want me to put these contraints in the paper?}
\todo[inline]{JS: Yes.}
\todo[inline]{KC: The machine definition gives a nice context for blame strategy so I 
define blame strategy later. You will find a todo near the blame strategy definition.}
%
The transitions involving casts are highlighted in red and described in
more detail below. The other transitions are standard for a CEK
machine for an extended simply typed lambda calculus~\citep{Felleisen:2009aa}.
%
Recall that CEK machines involve two kinds of transitions: those that
dive further into an expression (looking for a redex) and push an
entry onto the continuation and those that return a value to the
current continuation and possibly perform a computation.  That is, the
machine state is either in an evaluating $\sexpr{e}{E}{K}$ or
returning $\scont{v}{K}$ configuration, respectively. Additionally,
there are two halting configurations: \shalt{v} halts with a value,
while \error{l} halts with a blame label.  Observables include all the
value constructors and blame labels, as in the work of
\citet{siek2012interpretations}.  The function that converts values to
observables ($observe(v) = o$) is defined in the obvious way.

Let $v$ range over values. The $\vdyn{I}{v}$ stands for values casted
to the dynamic type, as in the work of \citet{wadler2009well}. The
inner value must be of an injectable type ($I$).
% The definition of
% injectable types depends on blame strategies:
The set of injectable type, which is defined by specify blame strategies,
must be a subset of pre-types.
%
For the \LD\ strategy, every pre-type is injectable 
(Fig.~\ref{fig:blame-strategy-LazyD});
%
for the \LUD\ strategy, only ground types are injectable 
(Fig.~\ref{fig:blame-strategy-LazyUD}).
%
Other kinds of values are either Boolean constants, functions, casted 
functions, pairs, or casted pairs.
%
The \proxy{v}{\ccast{S}{l}{T}} partial function applies the cast to $v$, which must
have type $S$. This function is handy when we describe the dynamic semantics.
% In earlier version of this work, 
% we were tempted to use a generic constructor for all casted value, and use 
% value typing rules to restrict the cast such that its source and target are 
% shallowly consistent pre-types. This approach unfortunately considers 
% $\ecast{\etrue}{\ccast{\Pbool}{l}{\Pbool}}$ as a value.

The syntax \error{l} occurs several times in the syntax definition. We shall 
view those nonterminals that have an \error{l} case as error monads. We use 
the symbol \mbind{} as the monad bind operator, which is defined as follows:
\[
\begin{array}{rcll}
(\error{l}) \mbind f & = & \error{l} \\
          a \mbind f & = & f(a)      & \text{otherwise}
\end{array}
\]
\todo[inline]{JS: why not use monadic let instead? That avoids
  the need to also use meta-lambdas, which you would need to
  explain, e.g., in the transition rule that uses applyCast.}
\todo[inline]{KC: 1. I will try to switch to monadic let. As you pointed out
when we wrote the WGT paper, let is more friendly to readers who are not familiar
with monads.
2. Do you imply that monadic let doesn't need to be explained?
}

@@

The transition relation \reduceto{s}{s} invokes the cast application
function, noted as $\applyCast{v}{\ccast{S}{l}{T}}$, which is also defined
by specify blame strategies.
% to capture the
% differences between blame strategies
% (Fig.~\ref{fig:blame-strategy-LazyD} and
% Fig.~\ref{fig:blame-strategy-LazyUD}).
%
When evaluating a cast expression, the machine moves the cast to the
continuation and evaluates the inner expression.
%
To cast a value, the machine invokes applyCast on the value. The cast
result is either a value or an error (\error{l}). If the cast
succeeds, the machine returns the value to the next continuation.  If
the cast fails, the machine halts with the blame label.
%
To apply a casted function, the machine first casts the argument $v$, then
applies the inner function $u$ to the casted argument, and finally
casts the return value.
%
To project the first element of a casted pair, the machine projects
the first element of the inner pair $u$ and cast it.
%
Projecting the second element of a casted pair is similar.

@@

\todo[inline]{KC: conditions of blame strategy}
\begin{definition}[Blame Strategy, $B$]
A blame strategy defines a subset of pre-types, written I, that is injectable,
and a function, written applyCast, that applies cast to values.
\end{definition}

@@

In this section, we review the coercions
of~\citet{henglein1994dynamic} and \citet{herman2010space} and the
normal form of \citet{siek2012interpretations} to motivate the design
of hypercoercions.
% We omit sum types and product types in this
% section because the prior work did not include them.

@@

Coercions come with a reduction relation so it is natural to ask about
their normal forms. \figref{fig:normal-coercion} defines the syntax of
coercion in normal form, written $s$ and $t$.
\todo[inline]{JS: In the figure, normal forms are written $s$ and $t$.}
\todo[inline]{KC: you are absolutely right.}
Conceptually, a coercion
in normal form has three parts, all of which are optional.  It may
start with a projection $I?^l$, followed by a function coercion, and
then concluded with an injection $I!$ or a failure $\bot^l$. While
this is a simple idea, 
% there are 10 clauses in the definition for $s$!
the three-part and symmetric structure is not obvious in the syntax definition.

@@

% All cast representation shall be proved correct, that is, besides reducing 
% space consumption, using a cast representation should not affect the results of 
% program execution.
In this section we present a framework for proving the 
correctness of cast representations. The framework includes an abstract data 
type named Cast ADT, which specifies the operators of cast representations, 
a CEK machine $\SMachine{C}$ parameterized over a Cast ADT,
which uses the cast operators implemented by $C$ to work with casts, 
and theorems relating $\SMachine{C}$ to $\CMachine{B}$.
%
Subsection~\ref{sec:cast-adt-and-space-efficient-cek} describes Cast ADT and the 
machine \SMachine{C}. We conjecture that the machine \SMachine{C} is 
space-efficient provided that the cast representation $C$ performs compression.
%
Subsection~\ref{sec:good-cast-adt} defines two subset of Cast ADT, namely Lazy D 
Cast ADT and Lazy UD Cast ADT. As their names suggest, if $C$ is a Lazy D Cast 
ADT, executing a program in $\SMachine{C}$ would give the same result as in 
$\LDMachine$. Similarly, if $C$ is a Lazy UD Cast ADT, executing a program in 
$\SMachine{C}$ would give the same result as in $\LUDMachine$.
%
We prove the theorems relating \SMachine{C} and \CMachine{B} in
Subsection~\ref{sec:S-correct}.
The framework is written in the Agda language~\citep{norell2008dependently}.
